\cchapter{Program Control}{program_control}
\label{chap:program_control}

Basic concepts and mechanisms for directing and controlling a program compilation and execution
are provided in this introduction and illustrated in subsequent examples.

\bigskip
CONDITIONAL COMPILATION and EXECUTION

Conditional compilation can be performed with conventional \bcode{\#ifdef} directives
in C, C++, and Fortran, and additionally with OpenMP sentinel (\scode{!$}) in Fortran. 
The \kcode{if} clause on some directives
can direct the runtime to ignore or alter the behavior of the construct.
Of course, the base-language \bcode{if} statements can be used to control the execution
of stand-alone directives (such as \kcode{flush}, \kcode{barrier}, \kcode{taskwait}, 
and  \kcode{taskyield}).
However, the directives must appear in a block structure, and not as a substatement.
The \kcode{metadirective} and \kcode{declare variant} directives provide conditional 
selection of directives and routines for compilation (and use), respectively.
The \kcode{assume} and \kcode{requires} directives provide invariants
for optimizing compilation, and essential features for compilation 
and correct execution, respectively.


\bigskip
CANCELLATION

Cancellation (termination) of the normal sequence of execution for the threads in an OpenMP region can
be  accomplished with the \kcode{cancel} construct.  The construct uses a
\plc{construct-type-clause} to set the region-type to activate for the cancellation. 
That is, inclusion  of one of the \plc{construct-type-clause} names \kcode{parallel}, \kcode{for}, 
\kcode{do}, \kcode{sections} or \kcode{taskgroup} on the directive line 
activates the corresponding region.  
The \kcode{cancel} construct is activated by the first encountering thread,  and it
continues execution at the end of the named region.
The \kcode{cancel} construct is also a cancellation point for any other thread of the team 
to also continue execution at the end of the named region.  

Also, once the specified region has been activated for cancellation any thread that encounters
a \kcode{cancellation point} construct with the same named region (\plc{construct-type-clause}),
continues execution at the end of the region.

For an activated \kcode{cancel taskgroup} construct, the tasks that
belong to the taskgroup set of the innermost enclosing taskgroup region will be canceled. 

A task that encounters a \kcode{cancel taskgroup} construct continues execution at the end of its
task region. Any task of the taskgroup that has already begun execution will run to completion,
unless it encounters a \kcode{cancellation point}; tasks that have not begun execution may be
discarded as completed tasks.

\pagebreak
CONTROL VARIABLES 

  Internal control variables (ICV) are used by implementations to hold values which control the execution
  of OpenMP regions.  Control (and hence the ICVs) may be set as implementation defaults, 
  or set and adjusted through environment variables, clauses, and API functions.  
 %Many of the ICV control values are accessible through API function calls.  
  Initial ICV values are reported by the runtime
  if the \kcode{OMP_DISPLAY_ENV} environment variable has been set to \vcode{TRUE} or \vcode{VERBOSE}. 

 %As an example, the \plc{nthreads-var} is the ICV that holds the number of threads
 %to be used in a \code{parallel} region.  It can be set with the \code{OMP\_NUM\_THREADS} environment variable, 
 %the \code{omp\_set\_num\_threads()} API function, or the \code{num\_threads} clause.  The default \plc{nthreads-var}
 %value is implementation defined.  All of the ICVs are presented in the \plc{Internal Control Variables} section
 %of the \plc{Directives} chapter of the OpenMP Specifications document.  Within the same document section, override 
 %relationships and scoping information can be found for applying user specifications and understanding the 
 %extent of the control.

\bigskip
NESTED CONSTRUCTS

Certain combinations of nested constructs are permitted, giving rise to \plc{combined} constructs
consisting of two or more directives.  These can be used when the two (or several) constructs would be used
immediately in succession (closely nested). A combined construct can use the clauses of the component
constructs without restrictions.
A \plc{composite} construct is a combined construct which has one or more clauses with (an often obviously)
modified or restricted meaning, relative to when the constructs are uncombined. %%[appear separately (singly).

%The combined \code{parallel do} and \code{parallel for} constructs are formed by combining the \code{parallel}
%construct with one of the loops constructs \code{do} or \code{for}.  The
%\code{parallel do SIMD} and \code{parallel for SIMD} constructs are composite constructs (composed from
%the parallel loop constructs and the \code{SIMD} construct), because the \code{collapse} clause must
%explicitly address the ordering of loop chunking \plc{and} SIMD ``combined'' execution.

Certain nestings are forbidden, and often the reasoning is obvious.  For example, worksharing constructs cannot be nested, and
the \kcode{barrier} construct cannot be nested inside a worksharing construct, or a \kcode{critical} construct. 
Also, \kcode{target} constructs cannot be nested, unless the nested target is a reverse offload.

The \kcode{parallel} construct can be nested, as well as the \kcode{task} construct.  
The parallel execution in the nested \kcode{parallel} construct(s) is controlled by the 
\kcode{OMP_MAX_ACTIVE_LEVELS} environment variable, and the \kcode{omp_set_max_active_levels} routine. 
Use the \kcode{omp_get_max_active_levels} routine to determine the maximum levels provided by an implementation.
As of OpenMP 5.0, use of the \kcode{OMP_NESTED} environment variable and the \kcode{omp_set_nested} routine 
has been deprecated.

More details on nesting can be found in the \docref{Nesting of Regions} of the \docref{Directives} 
chapter in the OpenMP Specifications document.


%===== Examples Sections =====
\input{program_control/assumption}
\input{program_control/cond_comp}
\input{program_control/icv}
\input{program_control/standalone}
\input{program_control/cancellation}
\input{program_control/requires}
\input{program_control/context_based_variants}
\input{program_control/dispatch}
\input{program_control/nested_loop}
\input{program_control/nesting_restrict}
\input{program_control/target_offload}
\input{program_control/pause_resource}
\input{program_control/reproducible}
\input{program_control/interop}
\input{program_control/utilities}

