\chapter*{Foreword}
\label{chap:foreword}

The OpenMP Examples document has been updated with new features
found in the OpenMP \PVER\ Specification. The additional examples and updates
are referenced in the Document Revision History of the Appendix on page~\pageref{chap:history}.

Text describing an example with a \PVER\ feature specifically states
that the feature support begins in the OpenMP \PVER\ Specification.  Also,
an \code{\small omp\_\PVER} keyword is included in the  metadata of the source code.
These distinctions are presented to remind readers that a \PVER\ compliant 
OpenMP implementation is necessary to use these features in codes.

Examples for most of the \PVER\ features are included in this document,
and incremental releases will become available as more feature examples
and updates are submitted and approved by the OpenMP Examples Subcommittee.

Examples are accepted for this document after discussions, revisions and reviews
in the Examples Subcommittee, and two reviews/discussions and two votes
in the OpenMP Language Committee.
Draft examples are often derived from case studies for new features in the language,
and are revised to illustrate the basic application of the features with code comments,
and a text description. We are grateful to the numerous members of the Language Committee 
who took the time to prepare codes and descriptions, and shepherd them through
the acceptance process. We sincerely appreciate the Example Subcommittee members, who 
actively participated and contributed in weekly meetings over the years.

\bigskip
Examples Subcommittee Co-chairs: \smallskip\linebreak
Henry Jin (\textsc{NASA} Ames Research Center) \linebreak
Swaroop Pophale (Oak Ridge National Labortory)


